---
title: "Analyse d'expression différentielle mutli-facteurs : Fer, Azote et CO2"
subtitle: "Clustering et réseau pour les gènes répondant à chaque facteur"
author: "Océane Cassan"
date: "19/03/2020"
output: 
  rmdformats::material:
    fig_width: 12
    highlight: kate

---

On cherche à étudier comparer des méthode d'inférence de réseaux (co-expression et régulation). 

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr, warn.conflicts = F, quietly = T)
library(rmdformats, warn.conflicts = F, quietly = T)

## Global options
options(max.print="75")
opts_chunk$set(cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               fig.align = "center")
opts_knit$set(width=75)
library(stringr)
library(GENIE3)
library(igraph)
library(visNetwork)
library(bc3net)

setwd("D:/These/NetworkInference")

load("./Data/DEGsListsFiltered.RData")
load("Data/PlnTFBDRegulatorsList.RData")
load("./Data/normalized.count_At.RData")
load("Data/OntologyAllGenes.RData")


source("Funtions/Network_functions.R")

genes <- DEGs[["cnF CnF"]]
rownames(ontologies) <- ontologies$ensembl_gene_id

removeNitrateStarv = F; removeIronStarv = T
 
 
if (removeIronStarv) normalized.count <- normalized.count[,grepl("F", colnames(normalized.count))]
if (removeNitrateStarv) normalized.count <- normalized.count[,grepl("N", colnames(normalized.count))]
  
```


# Personalized functions


```{r}
genie <- function(normalized.count, regressors=NA, targets=NA, nTrees=1000, nCores=5, top = 0.05, fixedLinkNumber = NA,
                  returnLinks = F){

  mat <- GENIE3(normalized.count, regulators = intersect(rownames(normalized.count),regressors), targets = targets, treeMethod = "RF", K = "sqrt", nTrees = nTrees, nCores = nCores,verbose = T)
  #hist(mat)
  links <- getLinkList(mat)
  
  if(returnLinks){return(links)}
  
  if(is.na(fixedLinkNumber)) links <- links[1:round(dim(links)[1]*top,0),]
  else links <- links[1:fixedLinkNumber,]
  
  print(paste0("Number of links : ", dim(links)[1]))
  
  g <- graph.data.frame(links, directed = F)
  return(g)
  
}


plotStats <- function(net){
  ont <- ontologies[match(V(net)$name, ontologies$ensembl_gene_id),]
  data <- toVisNetworkData(net)
  
  #attributs des noeuds et des liens
  data$nodes$Ontology <- ont[match(data$nodes$id, ont$ensembl_gene_id),]$external_gene_name
  data$nodes$description <-ont[match(data$nodes$id, ont$ensembl_gene_id),]$description
  data$nodes$group <- ifelse(data$nodes$id %in% TF$AGI, "Regulator", "Target Gene")
  
  importance <- (degree(net)/max(degree(net))+betweenness(net)/max(betweenness(net)))*0.5
  
  data$nodes$ranking <- importance[match(data$nodes$id, names(importance))]
  data$edges$value <- data$edges$weight
  netStats(net)
  #plot
  net <- visNetwork(nodes = data$nodes, edges = data$edges)%>% 
          visEdges(smooth = FALSE, color = '#333366') %>% 
          #visEdges(smooth = FALSE, arrows = 'to', color = '#333366') %>% 
          visPhysics(solver = "forceAtlas2Based", timestep = 0.9, minVelocity=10, 
                     maxVelocity = 10, stabilization = F)%>%
          visOptions(selectedBy = "group", highlightNearest = TRUE,nodesIdSelection  = TRUE, collapse = TRUE)%>% 
          visEvents(click = "function(nodes){
                  Shiny.onInputChange('click', nodes.nodes);
                  ;}") %>% 
          visGroups(groupname = "Regulator", size = 28,
                    color = list("background" = "#003399", "border"="#FFFFCC"), shape = "square") %>% 
          visGroups(groupname = "Target Gene", color = "#77EEAA") %>% 
          visNodes(borderWidth=0.5) 

  print(kable(head(data$nodes[order(-data$nodes$ranking),], n=20)))
  
  return(net)
}
```


# Know regulators as regressors

```{r}

plotStats(genie(normalized.count, regressors = intersect(genes, TF$AGI), targets = genes, top=1))

```


# All genes as regressors

```{r}

plotStats(genie(normalized.count, regressors = genes, targets = genes, fixedLinkNumber = 600))

```


# Thresold on links number effect


```{r}

links <- genie(normalized.count, targets = genes, regressors = intersect(genes, TF$AGI), returnLinks = T)
for(nbLinks in c(10, 100, 500, 1000, 2000)){
  newLinks <- links[1:nbLinks,]
  g <- graph.data.frame(newLinks, directed = F)
  V(g)$color <- ifelse(V(g)$name %in% TF$AGI, 1, 0)
  plot.igraph(g, vertex.size=4, vertex.label.cex=0.1, color = V(g)$color, main = paste0(nbLinks, " links"))
}
```


# GENIE3 VS BC3NET

The basic idea of BC3NET is to generate from one dataset D_s, consisting of s samples, an ensem-ble of B independent bootstrap datasets D_k by sampling from $D(s)$ with replacement by using anon-parametric bootstrap (Efron 1993).  Then, for each generated data set $D_k$ in the ensemble, anetwork $G^b_k$ is inferred by using C3NET (Altay 2010a). From the ensemble of networks $G^b_k$ we construct one weighted network $G^b_w$ which is used to determine the statistical significanceof the connection between gene pairs. This results in the final binary, undirected network G.

A base component of BC3NET is the inference method C3NET introduced in Altay (2010a), which we present in the following in a modified form to obtain a more efficient implementation.  Briefly,C3NET consists of three main steps.   First,  mutual information values among all gene pairs areestimated. Second, an extremal selection strategy is applied allowing each of the p genes in a givendataset to contribute at most one edge to the inferred network.  That means we need to test only pdifferent hypotheses and not $p(p-1)/2$.  This potential edge corresponds to the hypothesis test thatneeds to be conducted for each of the p genes.  Third, a multiple testing procedure is applied tocontrol the type one error. In the above described context, this results in a network $G^b_k$.


Avec GENIE3, il s'agit de méthodes ensemblistes.

```{r}

BC3net <- bc3net(dataset = normalized.count[genes,] )
gcc <- getgcc(BC3net)
plotStats(gcc)

genie <- genie(normalized.count, regressors = genes, targets = genes, fixedLinkNumber = length(E(gcc)))
plotStats(genie)

base::intersect(as.character(E(gcc)), as.character(E(genie)))

a <- E(gcc)
igraph::intersection(E(gcc)[1:length(E(gcc))], E(genie))

inter <- igraph::intersection(gcc, genie, keep.all.vertices=F)
plotStats(inter)


#################### Generation des reseaux : 

net <- genie(normalized.count, regressors = intersect(genes, TF$AGI), targets = genes, top=0.05)
data <- toVisNetworkData(net)
hist(data$edges$weight)
edges <- data$edges[,c("from", "to", "weight")]
colnames(edges)[1:2] <- c("src", "trg")
write.table(edges, file = "D:/These/Backboning/NetworkCO2_N_top0.05.txt", row.names = F, sep = '\t')


#### Reseau avec variance de 1 pour chaque gene
library(matrixStats)
dim(normalized.count)

normalized.count <- normalized.count[genes,]
normStd <- normalized.count/rowSds(normalized.count)
rowVars(normStd)

net <- genie(normalized.count, regressors = intersect(genes, TF$AGI), targets = genes, fixedLinkNumber=3000)
data <- toVisNetworkData(net)
data$nodes$Ontology <- ontologies[match(data$nodes$id, ontologies$ensembl_gene_id), "external_gene_name"]
data$nodes$description <- ontologies[match(data$nodes$id, ontologies$ensembl_gene_id), "description"]
data$nodes$group <- ifelse(data$nodes$id %in% TF$AGI, "Regulator", "Target Gene")


save(data, file = "D:/These/NetworkShiny/NetworkData/CO2DEGenes_faibleNitrate_CO2-N_VarianceNorm.RData")
hist(data$edges$weight)

################### generation pour tests avec une paire de TFs très corrélés

regressors = intersect(genes, TF$AGI)
# sans WRKY
regressors = regressors[regressors != "AT3G15500"]

net <- genie(normalized.count, regressors = regressors, targets = genes, fixedLinkNumber=3000)


data_sansWRKY <- toVisNetworkData(net)
data$nodes$Ontology <- ontologies[match(data$nodes$id, ontologies$ensembl_gene_id), "external_gene_name"]
data$nodes$description <- ontologies[match(data$nodes$id, ontologies$ensembl_gene_id), "description"]
data$nodes$group <- ifelse(data$nodes$id %in% TF$AGI, "Regulator", "Target Gene")


```

Ah.... ben c'est pas beaucoup en commun...... :(

# Réseau réponse nitrate


genes <- DEGs[["cnF CnF"]]

